Comparing probabilistic predictive models applied to football

### 2015-07-08 template.tex
revisao Ernesto

	In the literature, statistical models devoted to the prediction of
	football matches can be divided in two classes: models that make
	assumptions about the number of goals each team scores
	\citep{Maher82, Dixon97, Karlis2003} and models that make
	assumptions about the categorical final outcome (win, draw or defeat
	of the home team) \citep{Forrest2000, Koning2000, Brillinger2008,
		Brillinger2009}. For a discussion of these two approaches see
	\cite{Goddard2005}. The benchmark models considered in this work are
	part of the first approach, while the models proposed by us are part
	of the second one.

@@

@@

@@

>>>>

    In the literature, statistical models devoted to the prediction of
    football matches can be divided in two classes: models that make
    assumptions about the number of goals each team scores
    \citep{Maher82, Dixon97, Lee97, Karlis2003} and models that make
    assumptions about the categorical final outcome (win, draw or defeat
    of the home team) \citep{Forrest2000, Koning2000, Brillinger2008,
        Brillinger2009}. For a discussion of these two approaches see
    \cite{Goddard2005}. The benchmark models considered in this work are
    part of the first approach, while the models proposed by us are part
    of the second one.

@@

    The benchmark models assume that the number of goals $(Y_1, Y2)$ scored respectively by teams A (home team) and B (away
    team) follows a bivariate Poisson distribution \citep{Holgate64} with parameters $(\lambda_1, \lambda_2, \lambda_3)$,
    which has p.m.f given by
\begin{equation*}
P(Y_1 = y_1, Y_2 = y_2 | \lambda_1, \lambda_2, \lambda_3) =
\exp\{-(\lambda_1 + \lambda_2 + \lambda_3)\}
 \sum_{k = 0}^{\min(y_1, y_2)} \dfrac{\lambda_1^{x - k} \lambda_2^{y - k} \lambda_3^k}{(x-k)!(y -
 k)!k!}, \label{eq::pois.biv}
\end{equation*}
for $\lambda_1, \lambda_2 > 0$ and $\lambda_3 \geq 0$.

Both marginal distributions of $(Y_1, Y_2)$ follow Poisson
distributions with dependence parameter $\lambda_3 \geq 0$. If
$\lambda_3 = 0$ the marginal distributions are independent, while if
$\lambda_3 > 0$ the marginal distributions are positively
correlated. Because of its flexibility, \cite{Karlis2003} argues
that this distribution is a plausible choice for modeling dependence
of scores in sports competitions.

%Following the literature on soccer prediction \citep{Lee97, Karlis2003},
%both benchmark models assume log-linear link functions for $\lambda_1, \lambda_2$ and $\lambda_3$
%expressed in terms of team's attack and defensive abilities, and home advantage effect.

Similar to \cite{Lee97} and \cite{Karlis2003}, both benchmark models
adopt the following log-linear link functions
\begin{align*}
\log(\lambda_1) &= \mu + \text{ATT}_A - \text{DEF}_B + \gamma, \\
\log(\lambda_2) &= \mu + \text{ATT}_B - \text{DEF}_A,
\end{align*}
where $\mu$ is a parameter representing the average number of goals
in a match, $\text{ATT}_k$ is the offensive strength of team k,
$\text{DEF}_k$ is the defensive strength of team k and $\gamma$ is
the home advantage for team A, $k = A, B$.

The prediction for an upcoming match-day is obtained by fitting the
model to all the observed scores up to the previous match-day and
then summing up the probabilities of all scores relevant to the win,
draw and loss outcomes of future matches.

@@

        \bibitem{Holgate64}
        Holgate, P. Estimation for the bivariate Poisson distribution. {\em Biometrika} {\bf 1964}, {\em 51}, 241-287.

@@

        \bibitem{Lee97}
        Lee, A. J. Modeling scores in the Premier League: Is Manchester United really the best? {\em Chance} {\bf 1997}, {\em 10}, 15-19.
